Homework 5 - Stat 495

### 2022-10-19 homework/hmk5/hmk5Stat495F22.Rmd
Added intro and conclusion. Just need to edit

require(gridExtra)

library(ISLR)
library(glmnet)

@@

# Introduction

@@

Additionally, from the Scatterplot of Bedrooms and Price, there appears to be an outlier so we will remove that observation.

@@

fig6 <- ggplot(data = kchouse, aes(x = sqft_living, y = sqft_living15)) + 
  geom_point() + 
  labs(title = "Scatterplot of SqFt Living and SqFt Living in 2015", 
       x = "Sq Ft Living", y = "SqFt Living in 2015")
```

@@

From the scatterplot below of the latitude and the longitude, it seems like the houses are all bunched together in near each other (near Seattle) so the results of this analysis can only be generalized to that region of the state.

```{r}
ggplot(data = kchouse, aes(x = lat, y = long)) +
  geom_point()
```

@@

Next, since we want to know if the price of the house is greater or less than \$500,000, we will transform the price variable so that it is an binary predictor that is `1` if the price of the house is greater than \$500,000 and `0` if the price of the house is less than \$500,000.

```{r}
kchouse1 <- mutate(kchouse1, price = ifelse(price > 500000, 1, 0))
```

The last thing that we will do before creating the models is create a training and test set. The training set is used to create a model and then the test set is used to see how the model performs in making predictions. We will randomly create a 75%/25% split for the training/test set.

@@

The first models that we are going to be fitting are GLMs (or Generalized Linear Models) which are extensions of ordinary linear regression (fitting a least-squares curve) to cases where the response variables are in an exponential family form. This can be response variables that are binomial, Poisson, gamma, or beta. Since we transformed `price` into a binary variable, it can now be modeled as a binomial distribution.

First, we are going to fit a logistic model using all of the variables that we have not removed from the data set. We use the training set to create the model.

@@

```{r}
price.rpart1 <- rpart(price ~ ., data = train, method = "class")
printcp(price.rpart1)
rpart.plot(price.rpart1)
```

@@

# Conclusion

>>>>

require(gridExtra) 
library(rpart.plot)
library(ISLR)
library(glmnet)

@@

# Introduction 
We are interested in determining the features that are predictive of homes that are selling for more than half a million dollars. We are only interested in whether the price is over or under \$500,000, so will transform the price into a binary response variable where the the price is either over \$500,000 or it isn't. Then because the response variable is binary, we will use logistic regression to create the best model that we can. We will also perform a non-parametric analysis using classification trees in order to help visualize what predictors are best indicators of a house selling for greater than \$500,000. Finally, we will be able to compare the models to help answer the real estate developer's questions.

@@

Additionally, from the Scatterplot of Bedrooms and Price below, there appears to be an outlier so we will remove that observation.

@@

fig6 <- ggplot(data = kchouse, aes(x = sqft_living, y = sqft_living15)) + 
  geom_point() + 
  labs(title = "Scatterplot of SqFt Living and SqFt Living 2015", 
       x = "Sq Ft Living", y = "SqFt Living in 2015")
```

@@

From the scatterplot below of the latitude and the longitude, it seems like the houses are all bunched together near each other (close to Seattle) so the results of this analysis can only be generalized to that region of the state.

```{r}
ggplot(data = kchouse, aes(x = lat, y = long)) +
  geom_point() + 
  labs(title = "Scatterplot of Latitude and Longitude", 
       x = "Latitude", y = "Longitude")
```

@@

Next, since we want to know if the price of the house is greater or less than \$500,000, we will transform the price variable so that it is an binary predictor that is `1` if the price of the house is greater than \$500,000 and `0` if the price of the house is less than \$500,000. It appears that there are 12,546 houses with price below \$500,000 and 9,050 houses with a price above \$500,000.

```{r}
kchouse1 <- mutate(kchouse1, price = ifelse(price > 500000, 1, 0))
tally(~price, data = kchouse1)
```

The last thing that we will do before creating the models is create a training and test set. The training set is used to create a model and then the test set is used to see how the model performs in making predictions. We will randomly create a 75/25 split for the training/test set.

@@

The first models that we are going to be fitting are GLMs (or Generalized Linear Models) which are extensions of ordinary linear regression (fitting a least-squares curve) to cases where the response variables are in an exponential family form. This can be response variables that are binomial, Poisson, gamma, or beta. Since we transformed `price` into a binary variable, it can now be modeled as a binomial distribution. The model that we use for a binary response variable is a logistic regession model.

First, we are going to fit the logistic model using all of the variables that we have not removed from the data set. We use the training set to create the model.

@@

```{r, fig.width = 6, fig.height = 5}
price.rpart1 <- rpart(price ~ ., data = train, method = "class")
printcp(price.rpart1)
rpart.plot(price.rpart1)
```

@@

# Conclusion
The goal of this analysis was to determine the variables that are most indicative of a house's price being greater than \$500,000. To accomplish this, we first created two logistic regression models and decided that the one with less predictors was better of the two due to its comparable prediction accuracy and simplicity. This model contained the predictors `bathrooms`, `grade`, `yr_built`, `lat`, and `sqft_living15` which means that for the logistic regression model, these variables contributed the most to determining if a house had a price greater than half a million dollars. Secondly, we created two classification trees which we used to predict if a price was greater than \$500,000 or not. Again, we chose the model that had a similar prediction accuracy and less predictors. This tree was also much smaller, and thus easier to interpret. The predictors used in this model are `grade`, `lat`, and `sqft_living15`. The classification tree models performed slightly better than the logistic regression trees, but they were all within a small range so they are comparable. Combining information from both of our final models, `grade`, `lat`, and `sqft_living15` appeared as significant predictors in both models so these varaibles especially, are the ones that are most predictive of a house with a price over \$500,000.
